---
title: "Linear Regression using QR Decomposition"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to Linear Regression using QR Decomposition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The `AdvanceRAssignment4` package implements linear regression using QR decomposition method. This vignette describes how to use the `linreg` function with the `iris` dataset to fit a linear regression model, inspect outputs, and generate diagnostic plots.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Fitting a Linear Regression Model

To fit a linear regression model using the linreg function, the iris dataset will be used.

```{r}
library(AdvanceRAssignment4)

# Fit the model
model <- linreg(Sepal.Length ~ Sepal.Width + Petal.Length, data = iris)
```

## Accessing Model Coefficients

You can access the model coefficients using the `coef()` method:

```{r}
# Coefficients
model$coef()
```

## Making Predictions

To obtain predictions from the model, use the `pred()` method:

```{r}
# Predictions
predictions <- model$pred()
```

## Residuals

The residuals of the model can be accessed using the `resid()` method:

```{r}
# Residuals
residuals <- model$resid()
```

## Model Summary

To get a summary of the model, use the `summary()` method:

```{r}
# Model summary
model$summary()
```

## Generating Diagnostic Plots

To visualize model diagnostics, use the `plot()` method:

```{r}
# Diagnostic plots
model$plot()
```

This vignette offers a detailed guide on how to utilize the linreg function with the iris dataset, showcasing the comprehensive features of the AdvanceRAssignment4 package.
