---
title: "flight_delays"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{flight_delays}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This vignette demonstrates predictive modeling of flight delays using Ridge Regression with different values of the regularization parameter, `lambda`, to find the optimal model for our dataset.

## Step 1: Data Preparation

```{r data-preparation}
library(AdvanceRAssignment4)
library(nycflights13)
library(dplyr)
library(caret)

# Prepare the data
flights_weather <- flights %>%
  left_join(weather, by = c("year", "month", "day", "hour", "origin")) %>%
  select(dep_delay, arr_delay, temp, humid, wind_speed, visib) %>%
  mutate(
    temp_humid = temp * humid,        
    wind_vis = wind_speed * visib     
  ) %>%
  filter(!is.na(dep_delay)) %>%       
  na.omit()                           
```

## Step 2: Data Splitting

```{r data-splitting}
set.seed(1)
train_index <- createDataPartition(flights_weather$dep_delay, p = 0.8, list = FALSE)
train_data <- flights_weather[train_index, ]
temp_data <- flights_weather[-train_index, ]

# Split temp_data into validation and test sets (15% validation, 5% test)
validation_index <- createDataPartition(temp_data$dep_delay, p = 0.75, list = FALSE)
validation_data <- temp_data[validation_index, ]
test_data <- temp_data[-validation_index, ]
```

## Step 3: Model Training with Different Lambda Values

We train models with different lambda values and calculate the Root Mean Squared Error (RMSE) for each on the validation set.

```{r model-training}
lambdas <- 10^seq(-5, 5, length = 100)
results <- data.frame(lambda = lambdas, RMSE = NA)

for (i in seq_along(lambdas)) {
  model <- ridgereg$new(dep_delay ~ arr_delay + temp + humid + wind_speed + visib + temp_humid + wind_vis, 
                        data = train_data, lambda = lambdas[i])
  
  coefficients <- model$unScaledCoefficients
  
  validation_matrix <- model.matrix(dep_delay ~ arr_delay + temp + humid + wind_speed + visib + temp_humid + wind_vis, 
                                    data = validation_data)
  
  predictions <- as.vector(validation_matrix %*% coefficients)
  rmse <- sqrt(mean((validation_data$dep_delay - predictions)^2, na.rm = TRUE))
  results$RMSE[i] <- rmse
}

# Show the RMSE values for different lambda values
interval <- 10  
results_subset <- results[seq(1, nrow(results), by = interval), ]
results_subset
```

## Plotting RMSE vs Lambda

```{r plot-rmse, fig.width=6, fig.height=4}
plot(results$lambda, results$RMSE, type = "b", log = "x", xlab = "Lambda", ylab = "RMSE",
     main = "RMSE vs Lambda for Ridge Regression")
```

## Step 4: Find Optimal Lambda and Test Set Evaluation

```{r optimal-lambda}
best_lambda <- results$lambda[which.min(results$RMSE)]
cat("Optimal lambda:", best_lambda, "
")

final_model <- ridgereg$new(dep_delay ~ arr_delay + temp + humid + wind_speed + visib + temp_humid + wind_vis, 
                            data = train_data, lambda = best_lambda)

coefficients <- final_model$unScaledCoefficients

test_matrix <- model.matrix(dep_delay ~ arr_delay + temp + humid + wind_speed + visib + temp_humid + wind_vis, 
                            data = test_data)

final_predictions <- as.vector(test_matrix %*% coefficients)
final_rmse <- sqrt(mean((test_data$dep_delay - final_predictions)^2, na.rm = TRUE))
cat("Final RMSE on test data with optimal lambda:", final_rmse, "
")
```

This vignette provides a step-by-step process for finding the optimal lambda for ridge regression and evaluating the model on a test dataset, concluding with a plot of RMSE values across different lambdas.
