---
title: "ridgereg"
output: pdf_document
vignette: >
  %\VignetteIndexEntry{ridgereg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(AdvanceRAssignment4)
data("BostonHousing", package = "mlbench")
library("caret")
```

# Data Understanding

Data captures Boston Housing prices along with multiple other features of the city. 
```{r dataUnderstanding, fig.height=7, fig.width=10}
print(sapply(BostonHousing, class))
featurePlot(sapply(BostonHousing[, -14],as.numeric),
            as.numeric(BostonHousing$medv),
            plot = "scatter", layout = c(3, 5))
```

# Data Split

The dataset is split into testing and training dataset capturing 20% and 80% of 
the data respectively. To ensure effective and un-biased training and evaluation 
of the model, we used "createDataPartition" from "Caret" package. This ensures 
the similar distribution of dependent is maintained across both the split 
partitions.

```{r testTainSplit}
trainIndex <- createDataPartition(as.numeric(BostonHousing$medv), 
                                  p = .8, 
                                  list = FALSE, 
                                  times = 1)
trainData<-sapply(BostonHousing[trainIndex,],as.numeric)
testData<-sapply(BostonHousing[-trainIndex,],as.numeric)
```
# Model Training

Using Linear Regression Fit using Feed forward method for covariate(s) selection.
Root Mean Squared Error (RMSE) is used to identify best model trained on 
training data.

```{r trainLinearReg_FeedForward}

summary(myModel<-train(y = as.numeric(trainData[,14]), 
                       x = as.data.frame(trainData[,-14]),
                       'leapForward'))
```
Since only 4 variables are tested, I would use tuneGrid to test all variables

```{r trainLinearReg_FeedForward_TuneGrid}
summary(myModel<-train(y = as.numeric(trainData[,14]), 
               x =trainData[,-14],
               'leapForward',metric ="RMSE",maximize = FALSE,
               # trControl = trainControl(method = "none"),
               verbose =FALSE,tuneGrid = expand.grid(.nvmax = seq(1, 13))))
```

# Model Performance

As observed from model iterations, the R-Square increases with increasing number
of covariates. RMSE follows a similar increasing trend with increase in number of
covariates. For this reason the algorithm recommends a model with all covariates.
This is known to be true always. 

However, if observed closely, one would see the rate of growth of RMSE as well 
as R-Square decrease over time with maximum gainsobserved during addition of 
first few covariates to form an elbow curve. This curve can helps us balance 
between model complexity and model accuracy if required. Probably a model with 
just 2 covariates (lstat & rm) or (percentage of lower status of the population 
& average number of rooms per dwelling) might be sufficient to meet our goals 
and provides a good balance between accuracy or RMSE and model complexity.

```{r trainLinearReg_FeedForward_performance}
myModel
```
# Evaluating model performce on Testing dataset
We observe a very low RMSE on testing dataset when compared to that on training
dataset, indicating over-fitting. 

Probably we need to either explore a more simpler model or use an alternative 
approach to modeling to ensure the trained model does not over fit on the 
training data and provides similar results when tested on new data or 
in-production.

```{r trainLinearReg_FeedForward_performance_Testing}
predictions = predict(myModel, newdata = testData[,-14])
RMSE(predictions,testData[,14])
```

# Model Training Using ridgereg()

```{r trainRidgeRegression}

customRidgeReg<-list(type = "Regression",
              library = "AdvanceRAssignment4")

customRidgeReg$parameters<-data.frame(parameter=c("lambda"),
                                      class =c("numeric"),
                                      label = c("lambda"))
  
myModelRidreg<-ridgereg$new(formula=reformulate(colnames(trainData[,c(1,2,3)]),"medv"),
                            data=as.data.frame(trainData),lambda=1)
myModelRidreg
```

